# Answer for the Assignment
## Information regarding the DataSet
It contains monthly returns for five value-weighted portfolios from October 1969 to September
2019. The portfolios are formed using all stocks traded at NYSE, NASDAQ, and AMEX.
A proxy for the level of operating profitability (OP) is computed for each stock. Then, the stocks are sorted according to OP and are split evenly into 5 groups (quintiles).

LO contains the lowest OP quintile stocks whereas portfolio HI contains the highest OP quintile stocks. Portfolios QNT2, QNT3, and QNT4 are formed in the obvious fashion using the remaining stocks. The provided file also contains market returns (MKT)2 and T-bill rates (RF). Note that all returns are expressed in percentage points.

### Code
```{r}
# setting working directory
setwd("~/Documents/Coursework/COEC371/problem_sets/MarkedAssignment")
# importing libraries
library(tidyverse)
library(readxl)
library(stats)
library(plotly)
library(ggthemes)
```
Importing data from VW_5_OP_Mkt_Rf_Monthly_196910_201909.xlsx
```{r}
data =read_excel('~/Documents/Coursework/COEC371/problem_sets/MarkedAssignment/VW_5_OP_Mkt_Rf_Monthly_196910_201909.xlsx')
# calculating the risk premium
data <- data %>% mutate(RP = (MKT) - RF)
# summary(data)
# str(data)
```

#### Part A
to perform Regression Analysis for all five OP portfolios.
regress portfolio excess returns on market excess
returns.

What are the alpha estimates for the 5 OP portfolios and the standard errors of these estimates?
```{r initial}
# attempt to regress 
reg1 = data %>% lm(formula = (LO-RF) ~ RP)
reg2 = data %>% lm(formula = (QNT2-RF) ~ RP)
reg3 = data %>% lm(formula = ((QNT3)-RF) ~ RP)
reg4 = data %>% lm(formula = ((QNT4)-RF) ~ RP)
reg5 = data %>% lm(formula = ((HI)-RF) ~ RP)

#create a table
stock <- c('LO','QNT2','QNT3','QNT4','HI')
betas <- c(reg1$coefficients[2],
        reg2$coefficients[2],
        reg3$coefficients[2],
        reg4$coefficients[2],
        reg5$coefficients[2])
alphas <- c(reg1$coefficients[1],
        reg2$coefficients[1],
        reg3$coefficients[1],
        reg4$coefficients[1],
        reg5$coefficients[1])
df <- tibble(stock,alphas,betas)
df
```
according to the betas above, LO should have the highest return and QNT3 should have the lowest.


#### Which portfolios are fairly priced, overpriced, or underpriced from a CAPM perspective? [Hint: For which portfolios are the alpha estimates statistically significant?]

*just to see how the data distribution looks*
I have commented this out, as its not helpful, please refer to the previous file, to see the density plots
```{r density plots}
# density plots
# plot(density(data$LO))
# plot(density(data$QNT2))
# plot(density(data$QNT3))
# plot(density(data$QNT4))
# plot(density(data$HI))
```
### Plot SML, as predicted by the CAPM
So essentially what i am doing here, is plotting the portfolio excess returns against the market excess returns, then i generate the SML by regressing both excess returns. (essentially you could think of it as a line of best fit)

```{r part f plot}
# create a annnualised data set
Annual_data <- data %>% select(LO,QNT2,QNT3,QNT4,HI,MKT,RF,RP)
Annual_data <- Annual_data*12

# plot for part f
plot1 <- Annual_data %>% ggplot( aes(y = (LO-RF), x = RP))+geom_point(col='yellow4') + xlab('Market Excess Return') + ylab('Portfolio Excess Return') + ggtitle('Does CAPM hold for every portfolio?') + geom_smooth(method='lm')+geom_point(aes(y = (QNT2-RF)),col='yellow3')+geom_point(aes(y=(QNT3-RF)),col='yellow2')+geom_point(aes(y=(QNT4-RF)),col='yellow1')+geom_point(aes(y=(HI-RF)),col='lightyellow')
plot1 <- ggplotly(plot1)
plot1
```

###### question
On a new graph, plot the *CAPM-predicted risk premium (y-axis) vs. the realized excess return (x-axis) for each portfolio*. Note that (i) the CAPM-predicted risk premium is given by beta times average market excess return and (ii) the realized risk premium is equal to average excess return. Use annualized returns for easier interpre- tation. What should the plot look like if the CAPM holds? What can you say about the validity of the CAPM from this graph? [Hint: plot the 45-degree line on the same graph.]

###### answer
I have plotted the points, as beta times the annualized market excess return, I removed the 45 degree line, cause it made sense as you look at it.

```{r part g plot *need to add 45 degree line*}
# regression using annualised data
a_reg1 = Annual_data %>% lm(formula = (LO-RF) ~ RP)
a_reg2 = Annual_data %>% lm(formula = (QNT2-RF) ~ RP)
a_reg3 = Annual_data %>% lm(formula = ((QNT3)-RF) ~ RP)
a_reg4 = Annual_data %>% lm(formula = ((QNT4)-RF) ~ RP)
a_reg5 = Annual_data %>% lm(formula = ((HI)-RF) ~ RP)

#plot 
CAPM <- Annual_data %>% ggplot( aes(y = (a_reg1$coefficients[2]*mean(RP)), x = mean(LO)))+geom_point(col='yellow4')+xlab('Realised Excess Return') + ylab('CAPM - predicted risk premium') + ggtitle('Does CAPM hold for every portfolio?') +geom_point(aes(y = (a_reg2$coefficients[2]*mean(RP)), x = mean(QNT2)),col='yellow3')+geom_point(aes(y = (a_reg3$coefficients[2]*mean(RP)), x = mean(QNT3)),col='yellow2')+geom_point(aes(y = (a_reg4$coefficients[2]*mean(RP)), x = mean(QNT4)),col='yellow1')+geom_point(aes(y = (a_reg5$coefficients[2]*mean(RP)), x = mean(HI)),col='lightyellow')
CAPM <- ggplotly(CAPM)
CAPM
```

#### PART B
initial data transformation, commented out some extra stuff
```{r part a}
# expected values
E_LO = mean(Annual_data$LO)
E_HI = mean(Annual_data$HI)
E_RF = mean(Annual_data$RF)

# risk
# look into why  using the annuaalised daataset dosent work here.

sd_LO = sd(data$LO)
sd_HI = sd(data$HI)

# correlation
HILO = cor(Annual_data$LO,Annual_data$HI)

#table
answer = tibble(E_LO,E_HI,sd_LO,sd_HI,HILO)
#answer
```
```{r part b}
annualised_er = answer[1:2]
annualised_sd = answer[3:4]* sqrt(12)

annualised_er
annualised_sd
#correlation 
#HILO
```
```{r Sharpe Ratio}
# Sharpe ratio calculation
SharpeRatio_LO = (annualised_er - 3)/annualised_sd
SharpeRatio_LO = rename(SharpeRatio_LO, SR_LO = E_LO,SR_HI = E_HI)
SharpeRatio_LO
```
 As can be seen from the output, HI provides higher return for the given risk.
```{r Portfolio}
# creating weights
weights <-seq(from =-1.5,to = 1.5,length.out=30)

# data table with the weights
df_w <- tibble(wLO = weights, 
               wHI = (1-weights))

# table with er and sd for the given wieghts 
B_df <- df_w %>% mutate(er_P = wLO*E_LO + wHI*E_HI,sd_P = sqrt((wLO*sd_LO*sqrt(12))^2 + (wHI*sd_HI*sqrt(12))^2 + 2*wLO*wHI*HILO))
#B_df
```

##### top 6 entries with maximum Portfolio risk

```{r 6_max_risk_portfolio}
# top 6 entries maximum Portfolio SD
head(arrange(B_df,desc(sd_P)))

```

## COMPLETE PORTFOLIO
**recheck part d**
```{r}
#weights
cp_weights <-seq(from =-1.5,to = 1.5,length.out=30)

# data table with weights
df_cp <- tibble(wRF = cp_weights, 
               wP = (1-cp_weights))
# table with cp
cp_df <- B_df %>% mutate(er_CP = df_cp$wRF*0.03 + df_cp$wP*er_P,sd_CP = (df_cp$wP*sd_P))
#cp_df

# plot
fml <- ggplot()+geom_point(data = B_df,aes(y=er_P,x=sd_P))+geom_point(data = tibble(sd =c(sd_LO*sqrt(12),sd_HI*sqrt(12)),mean = c(E_LO,E_HI)),aes(x=sd,y=mean),color = "red",size=3,shape=18)+ggtitle("Portfolio returns",subtitle = "with LO and HI")+xlab("standard deviation")+ylab("expected returns")+scale_y_continuous()+scale_x_continuous()+geom_hline(yintercept = 3)
fml <- ggplotly(fml)
fml
```

Personal note, need to look into, why the red line is the way it is.

### tangent portfolio weights

```{r}
numerator = ((mean(E_LO)-3)*sd_HI -(mean(E_HI)-3))*HILO
denominator = (mean(E_LO)-3)*sd_HI +(mean(E_HI)-3)*sd_LO-(mean(E_HI)-3+mean(E_LO)-3)*HILO
cw_LO = numerator/denominator
#cw_LO
cw_HI = 1-cw_LO
# lets compute the er and sd
er_cp = cw_LO*E_LO + cw_HI*E_HI
sd_cp = sqrt((cw_LO*sd_LO*sqrt(12))^2 + (cw_HI*sd_HI*sqrt(12))^2 + 2*cw_LO*cw_HI*HILO)
# sharpe ratio
sharpe_cp = (er_cp - 3)/sd_cp
```

##### Tangent Portfolio Table

The table below provides the Expected Return, Standard Deviation and the Sharpe Ratio of the Tangent Portfolio.

```{r table with Complete Portfolio stuff}
#output
tibble(er_cp,
sd_cp,
sharpe_cp)

```

now that we have formed *the tangent line*, and figured out the tangent portfolio. now we will try to find the complete portfolio, with 12% expected return. just a quick reminder, complete portfolio is any portfolio of risky assets adjoined with a risk free asset.

```{r last part}
# expected returns and the sd for the tangent portfolio
e_tp = df_cp$wP*er_cp + df_cp$wRF*3
sd_tp = df_cp$wP*sd_cp
final_df = tibble(e_tp,sd_tp,w_a = df_cp$wP*er_cp,w_LO = df_w$wLO,w_HI = df_w$wHI ,w_rf= df_cp$wRF*3)
max(final_df$e_tp)
```

#### Answer for part f
the table below shows the, 12.1 expected return and weights associated with them
```{r}
# the filter to obtain the portfolio with 12% returns
portfolio = filter(final_df,e_tp >= 12 & e_tp <=12.9)
portfolio
```
 
 
#### Plot below

```{r plotly plot}
test <- ggplot()+geom_point(data = B_df,aes(y=er_P,x=sd_P))+geom_point(data = tibble(sd =c(sd_LO*sqrt(12),sd_HI*sqrt(12)),mean = c(E_LO,E_HI)),aes(x=sd,y=mean),color = "red",size=3,shape=18)+ggtitle("Portfolio returns")+xlab("standard deviation")+ylab("expected returns")+scale_y_continuous()+scale_x_continuous()+geom_hline(yintercept = 3)+geom_line(data = final_df,aes(y=e_tp,x=sd_tp),color='blue')+geom_point(data = portfolio, aes(y = e_tp,x = sd_tp),color = "green",size=4,shape=18)
test <- ggplotly(test)
test
```
