In this file, I am going to attempt to solve:

3.  You consider investing in two stocks, A and B. The expected return on stock A is 11% and the standard deviation is 16%.  The expected return on stock B is 17% and thestandard deviation is 26%.
  (a)  The correlation between stocks A and B is 0.3.  Compute the expected return andstandard deviation of a portfolio that has 0% A, 10% A, 20% A, etc, until 100%A. Plot the portfolio frontier formed by these portfolios.
```{r}
# imported all the libraries
library(tidyverse)
library(ggthemes)
```

stock information
```{r}
#expected returns on both stocks
er_A=0.11
er_B=0.17

#standard deviation on both stocks
sd_A=0.16
sd_B=0.26

#correlation coefficient
ro = 0.3
new_ro = 0.8

#create weights
A_weights <-seq(from =0,to = 1,length.out=10)

#data table with weights
df_w <- tibble(wA=A_weights,
             wB=(1-A_weights))

#calculate portfolio returns and s.d.
df <- df_w %>% mutate(er_P = wA*er_A + wB*er_B,sd_P = sqrt((wA*sd_A)^2 + (wB*sd_B)^2 + 2*ro*wA*wB*sd_A*sd_B))
df

```
Plot 1
```{r}
#time to plot the data
ggplot() +
geom_point(data = df, aes(x = sd_P, y = er_P,)) +
geom_point(data = tibble(sd = c(sd_A, sd_B), mean = c(er_A, er_B)),
aes(x = sd, y = mean), color = "red", size = 3, shape = 18)+ggtitle("Portfolio returns",subtitle = "ro = 0.3") +xlab("standard deviation of the portfolio")+ylab("expected return of the portfolio")+scale_y_continuous()+scale_x_continuous()+theme_economist()
```
 (b)  Repeat the previous question, assuming that the correlation is 0.8.
```{r}
# calculate new portfolio returns and s.d.
new_df <- df_w %>% mutate(er_P = wA*er_A + wB*er_B,sd_P = sqrt((wA*sd_A)^2 + (wB*sd_B)^2 + 2*new_ro*wA*wB*sd_A*sd_B))
new_df

```
Plot2
```{r}
#plot 
ggplot()+geom_point(data = new_df,aes(y=er_P,x=sd_P))+geom_point(data = tibble(sd =c(sd_A,sd_B),mean = c(er_A,er_B)),aes(x=sd,y=mean),color = "red",size=3,shape=18)+ggtitle("Portfolio returns",subtitle = "ro = 0.8")+xlab("standard deviation")+ylab("expected returns")+scale_y_continuous()+scale_x_continuous()+theme_economist()
```
(c)  Explain intuitively why the portfolio frontier is different in the two cases.
since the correlation is high in plot2, both stocks move in the same pattern. Hence not many gains can be achieved through diversification, as a result the minimum variance is achieved by holding A (100%). On the other hand in plot 1, as the corrleation is low, the portoflio with positive weights on A and B can have a lower variance than holding both A and B.

References:
This work is really similar to what i have done here.
https://www.r-bloggers.com/a-gentle-introduction-to-finance-using-r-efficient-frontier-and-capm-part-1/